---
title: "ConcreteImpactApp: Life Cycle Assessment of SCM Concrete"
author:
  - name: "Duah Philip"
    affiliation: "Mine Sustainability Modeling Research Group, Missouri University of Science and Technology"
    email: "dpbxc@mst.edu"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 7
vignette: >
  %\VignetteIndexEntry{ConcreteImpactApp: Life Cycle Assessment of SCM Concrete}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center"
)
```

## Overview

`ConcreteImpactApp` is an R package providing an interactive Shiny dashboard
and a set of data-access functions for evaluating the cradle-to-gate
life cycle environmental impacts of U.S. ready-mix concrete produced with
supplementary cementitious materials (SCMs): fly ash (FA) and ground granulated
blast-furnace slag (GGBFS).

The tool implements the **TRACI 2.2** impact assessment methodology across
**10 environmental impact categories** and covers:

- **Fly ash**: 16 scenarios (3 treatment technologies × 3 source types × dose levels),
  with and without avoided-impact credit allocation
- **Slag (GGBFS)**: 4 global supply-chain origins (USA, Brazil, China, Japan)
- **6 compressive strength classes**: 17.2, 20.7, 27.6, 34.5, 41.4, 55.2 MPa

---

## Installation

```r
# Install from the bundled tar.gz (as distributed with the manuscript)
install.packages("ConcreteImpactApp_1.0.0.tar.gz",
                 repos = NULL, type = "source")

# Or install development version from GitHub
# remotes::install_github("Duah-Philip/ConcreteImpactApp")
```

Dependencies are automatically installed:

```r
# Core Shiny stack
library(shiny)
library(shinydashboard)

# Data 
library(readxl)

# Visualisation
library(ggplot2)
library(plotly)
library(DT)
```

---

## Launching the Interactive Dashboard

```r
library(ConcreteImpactApp)
run_app()
```

This opens the browser-based dashboard. No internet connection is required;
all LCIA data are bundled within the package.

---

##

The package also exposes its data-reading functions for scripted analyses,
making the underlying data fully reproducible outside the Shiny interface.

```{r load-pkg}
library(ConcreteImpactApp)
```

### Fly Ash — Absolute LCIA Values

```{r fa-lcia-abs}
# Read absolute LCIA values for Fresh Ash, Electrostatic 12% LOI, with credit
df_fa <- read_fa_lcia("FreshAsh_ET12", credited = TRUE)

# Filter to 27.6 MPa concrete
df_276 <- df_fa[abs(df_fa$Strength_MPa - 27.6) < 0.5, ]

# Display
knitr::kable(
  df_276[, 1:6],
  caption = "Absolute LCIA values — Fresh Ash ET 12%, 27.6 MPa (first 5 categories)",
  digits  = 4
)
```

### Fly Ash — Percent Change vs Baseline

```{r fa-pct-data}
# Read percent-change data (all strengths and replacement levels)
df_pct <- read_fa_pct_full("FreshAsh_ET12", credited = TRUE)

cat("Rows:", nrow(df_pct), "\n")
cat("Compressive strengths (MPa):", paste(sort(unique(df_pct$str_num)), collapse=", "), "\n")
cat("FA replacement levels (%):", paste(sort(unique(df_pct$FA_Pct)), collapse=", "), "\n")
cat("Impact categories:", length(unique(df_pct$Impact_Category)), "\n")
```

### Fly Ash — Facet Plot (all strengths and replacements)

```{r fa-pct-plot, fig.cap="Percent change in LCIA impact categories for Fresh Ash ET 12% scenario with avoided-impact credit. Bars represent fly ash replacement levels (20%, 30%, 40%) relative to the 0% Portland cement baseline. Green bars indicate environmental improvements; purple bars indicate increases.", fig.width=10, fig.height=7}
df_pct$FA_Pct <- as.character(df_pct$FA_Pct)

make_fa_pct_facet(
  df           = df_pct,
  title_txt    = "% Change vs Baseline — Fresh Ash | Electrostatic (ET) 12% LOI",
  subtitle_txt = "With Avoided-Impact Credit | All Compressive Strengths | FA 20%, 30%, 40%"
)
```

### Slag — Percent Change vs Baseline

```{r slag-pct-plot, fig.cap="Percent change in LCIA impact categories for USA domestic slag (GGBFS) supply chain. Bars represent GGBFS replacement levels (30%, 40%, 50%) relative to the 0% Portland cement baseline.", fig.width=10, fig.height=7}
df_slag <- read_slag_pct_full("USA")
df_slag$GGBFS_Pct <- as.character(df_slag$GGBFS_Pct)

make_slag_pct_facet(
  df           = df_slag,
  title_txt    = "% Change vs Baseline — Slag (GGBFS) | USA Domestic Supply Chain",
  subtitle_txt = "All Compressive Strengths | GGBFS 30%, 40%, 50%"
)
```

### Mix Design Reference Data

```{r mix-design}
# Fly ash mix designs
# Access internal mix design data via package namespace
FA_MIX <- ConcreteImpactApp:::FA_MIX

fa_27 <- FA_MIX[abs(FA_MIX$strength - 27.6) < 0.5,
                c("pct_fa", "cement", "fly_ash", "water")]
fa_27$pct_fa <- paste0(fa_27$pct_fa, "%")

knitr::kable(fa_27,
  col.names = c("FA Replacement", "Cement (kg/m3)", "Fly Ash (kg/m3)", "Water (kg/m3)"),
  caption   = "Fly ash concrete mix designs at 27.6 MPa (NRMCA 2021)",
  row.names = FALSE
)
```

---

## Data Sources and Methodology

See paper/Manuscript 

### Impact Categories (TRACI 2.2)

| Short Name              | Unit            |
|-------------------------|-----------------|
| Global Warming          | kg CO2 eq       |
| Ozone Depletion         | kg CFC-11 eq    |
| Smog                    | kg O3 eq        |
| Acidification           | kg SO2 eq       |
| Carcinogenics           | CTUh            |
| Non-Carcinogenics       | CTUh            |
| Respiratory Effects     | kg PM2.5 eq     |
| Ecotoxicity             | CTUe            |
| Freshwater Eutrophication | kg P eq       |
| Marine Eutrophication   | kg N eq         |

### Ash Processing Technologies

- **Thermal (TT)**: Combustion-based reduction of Loss on Ignition (LOI)
  at 6% and 12% carbon content
- **Electrostatic (ET)**: Electrostatic separation of unburned carbon
  particles at 6% and 12% LOI
- **Chemical (CHEM)**: Non-ionic surfactant treatment at a commonly used
  (Typical) dose and a Maximum dose

### Fly Ash Source Types

- **Fresh Ash**: Directly from a coal-fired power plant
- **Landfill**: Recovered from dry landfill disposal
- **Impoundment**: Recovered from wet ash pond (impoundment)

---

## Session Information

```{r session-info}
sessionInfo()
```
